---
title: Automate your CV/Resume writing process (using RMarkdown)
author: Priyanka
date: '2021-07-19'
slug: automate-your-cv-resume-writing-process-using-rmarkdown
categories:
  - R
tags:
  - R Markdown, CV, {vitae},{datadrivencv}
---


Credits first: 
Thanks @jonthegeek for this [article](http://jonthegeek.com/2018/02/27/blogging-in-ten-minutes/) and for the inspiration to get me started with my blog yay!! This is by far the best article I have related to achieve any data related tasks without making (almost) no mistakes/errors. 

Thanks also to Nick Strayer, Bryan Jenks and Mitchell O'Hara-Wild and all those involved with the 2 packages I am going to write about in the rest of the blog. 

Now to the actual post on how to create your cv/resume through an automated process: 

So when I started working on mine I came across 2 options to do this: 

1. [{vitae}](https://cran.r-project.org/web/packages/vitae/vignettes/vitae.html)

2. [{datadrivencv}](http://nickstrayer.me/datadrivencv/)

There were a few blog posts I could find using each. The idea is to have a central place where you want to add/edit your career updates only in a simple text format and not have to deal with formatting issues (which you would otherwise face, almost everytime if you are working on any word software).

With my first reading of both, I liked and started working through [Nick's article](https://livefreeordichotomize.com/2019/09/04/building_a_data_driven_cv_with_r). However, I ran into certain issues:    

1. the files I downloaded did not exactly match with what were used in the article (they must have been updated at some point, the github link has 86 commits after all).

2. the two .R files (CV_printing_functions.R, render_cv.R) had lots of code to go through to make changes to make it yours.

3. (main culprit why I quit) after i created  copy of my googlesheet, I ran into authentication issues which I could resolve if I run the code to read data file in console but not through the code functions. So that was a bummer, i made progress with inputs from wonderful R4DS community however since I was in rush of getting my automated CV out I moved on to trying the {vitae} approach.

So, here's a run-down of what I did to get to this automated resume on mine, available here for [sample](https://github.com/priyankagagneja/cv-automation-in-r/blob/master/Priyanka_Gagneja_CV.pdf) 

1. Download the package {vitae} e.g.:
```{r, eval=FALSE, include=TRUE}
install.packages('vitae')
```

2. Install tinytex as this [article](https://pkg.mitchelloharawild.com/vitae/) says. I had this already, so I moved to the next step (hoping that i would be up & running sooner than anybody else... duh!!).

3. I copied 2 files (bare minimum requirements I think!) from Bryan Jenks's CV repo [here](https://github.com/tallguyjenks/CV) or you could use mine [here](https://github.com/priyankagagneja/cv-automation-in-r) (i did not create a folder structure as I intentionally kept it minimal code) for a cleaner version - `CV.Rmd` and `r/data.R`

4. You could simply test by knitting the CV.Rmd as it is. 

5. You may find some knitting errors relating to LaTex. I had to install xelatex (Basic Tex for Mac) hence from [here](http://www.texts.io/support/0001/) and the links i used for it were: 

(a) Start with this command (source: [here](https://gist.github.com/peterhurford/75957ba9335e755013b87254ec85fab1))
```{r, eval=FALSE, include=TRUE}
sudo tlmgr update --self
```

(b) Followed by installing any package or .sty files errors you would get. For e.g.: 

```{r, eval=FALSE, include=TRUE}
sudo tlmgr install enumitem
```

(c) In some cases, the name of .sty file will not match with the package needed exactly, in which you would need to [identify the package](https://bookdown.org/yihui/rmarkdown-cookbook/install-latex-pkgs.html) using 
```{r, eval=FALSE, include=TRUE} 
# use the `text` argument
tinytex::parse_packages(
  text = "! LaTeX Error: File `ocgbase.sty' not found."   # replace ocgbase with whatever file is missing for you
)
# it should return "ocgx2"`
```

and then install that identified package in your terminal. 
```{r, eval=FALSE, include=TRUE}
sudo tlmgr install ocgx2
```

6. Once all of these errors are sorted, you are pretty much set. Edit the files `data.R` and `CV.Rmd` with your information. This setup mainly uses 2 functions `detailed_entries()` and `brief_entries()`. If you see my CV you may notice I played around with the what,where, why arguments a little bit to get the aesthetics I wanted for my CV. 

7. I did not have this step but for the academicians(Cath this is for you), there might be an additional step of using additional function `bibliography_entries()`. [Package documentation](https://pkg.mitchelloharawild.com/vitae/) also has examples on connecting with your orcid id such that it can always access the most recent state of your publications. 

When/If you try to change the output format, you may encounter some more missing .sty files and you can go back to step 5(c) and that should be it. 

I hope this was useful and less taxing for atleast some people here in R community. 
Happy to take any feedback/comments/questions!!

